Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-31 paper.qmd
working on correct sizing of figures

## Best fit

@@

error_plot_two_non_linear_diff_shaped_close_clusters <-
  plot_mse(error_two_non_linear_diff_shaped_close_clusters) +
  scale_x_continuous(breaks = sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[seq(1, length(unique(error_two_non_linear_diff_shaped_close_clusters$a1)), by = 5)]) +
  scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363', '#984ea3')) +
  scale_y_log10()

```{r}
#| fig-cap: "Assessing which of the 6 NLDR layouts on the two non-linear clusters data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). Layout d is universally poor. Layouts a, b, e that show two close clusters are universally suboptimal. Layout b with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases. Layout e has small separation with oddly shaped clusters. Layout a is the best choice."
#| label: fig-two_non_linear_diff_shaped_close_clusters-mse
#| fig-pos: H
#| out-width: 100%

free(error_plot_two_non_linear_diff_shaped_close_clusters) + wrap_plots(nldr_two_non_linear_diff_shaped_close_clusters1, nldr_two_non_linear_diff_shaped_close_clusters2, nldr_two_non_linear_diff_shaped_close_clusters3, 
                                    nldr_two_non_linear_diff_shaped_close_clusters4, nldr_two_non_linear_diff_shaped_close_clusters5,
                                    nldr_two_non_linear_diff_shaped_close_clusters6, ncol = 2)
```

>>>>

## Best fit, or at least avoiding an inaccurate representation

@fig-toy-mse shows a summary of plots that can be used to help assess the strength of the fit, and compare the fits for different representations. What does it mean to be a best fit for this problem? There probably isn't a best fit, but there are wrong representations. The goal is to help users decide on a useful and appropriate low-dimensional representation of the high-dimensional data.

@@

error_plot_two_non_linear_diff_shaped_close_clusters <-
  plot_mse(error_two_non_linear_diff_shaped_close_clusters) +
  scale_x_continuous(breaks =
    sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[
      seq(1, length(
        unique(error_two_non_linear_diff_shaped_close_clusters$a1)), 
        by = 5)]) +
  scale_color_manual(
    values=c('#e41a1c','#ff7f00','#4daf4a', 
             "#a65628",'#636363', '#984ea3')) +
  scale_y_log10() + 
  theme(aspect.ratio = 1.5)
```

```{r}
#| fig-cap: "Assessing which of the 6 NLDR layouts on the two non-linear clusters data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). Layout d is universally poor. Layouts a, b, e that show two close clusters are universally suboptimal. Layout b with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases. Layout e has small separation with oddly shaped clusters. Layout a is the best choice."
#| label: fig-toy-mse
#| fig-pos: H
#| out-width: 100%
#| fig-height: 6
#| fig-width: 6

error_plot_two_non_linear_diff_shaped_close_clusters + wrap_plots(nldr_two_non_linear_diff_shaped_close_clusters1, nldr_two_non_linear_diff_shaped_close_clusters2, nldr_two_non_linear_diff_shaped_close_clusters3, 
                                    nldr_two_non_linear_diff_shaped_close_clusters4, nldr_two_non_linear_diff_shaped_close_clusters5,
                                    nldr_two_non_linear_diff_shaped_close_clusters6, ncol = 2, 
                                    widths = c(50, 50))
```
